Molecular dynamics simulation of synchronization in driven particles

### 2020-12-24 manuscript/manuscript_ajp_2020.tex
updated josephson junction and colloid text

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to interaction or an external stimulus \cite{Pikovsky2003}.
Synchronous behaviors where oscillators move in-phase (i.e. in time)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016} 
and metronomes coupled through the the supporting table \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation such as when 
birds coordinate
flapping of wings in order to optimize energy use during flight \cite{Portugal2014},
frogs communicate by croaking patterns modulated by location \cite{Aihara2014},
and humans synchronize clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two pendula on wall-mounted clocks.
Huygens demonstrated the tendency of 
two periodic oscillators in close
proximity to swing in time %, i.e. to synchronize,
due to interactions through the wall 
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different frequencies couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986} .
This can be demonstrated
with Lissajous figures, i.e. parametric graphs of two periodic functions
which make looped patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations

Synchronization is often achieved via external forcing
such as when
%single particle oscillators
pushing a child on a swing at the high point on her path, or
using pacemaker (an electrical generator)
pulses to regulate a heart beat.
%multiple particle oscillators
[Fireflies and pulsed LED].

Dynamical mode locking 
has many important
technological applications
and is often studied with simplified computational models.
These studies can be performed
with coupled oscillators \cite{Kuramoto193?}
or with a single particle driven across a periodic potential landscape \cite{}.
%Single particle models avoid many-body effects
%while offering insight.

%Oscillators ordered into chains or lattices.
Synchronization 
is %often
observed
in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}
i.e. two superconductors sandwiching an insulating layer.
When subject to an external voltage,
Cooper pairs in the superconducting materials
tunnel through the insulating layer.
Phase-locking is observed as 
%phase-locked current loops \cite{} 
stepped regions in current-voltage (I-V) relationship in these devices,
where a constant current results from a range of voltages
due to the quantized nature of the tunneling probability [NOT QUITE RIGHT...].
%instead of ohmic relationships.
Known as Shapiro steps \cite{Shapiro1963},
these indicators of phase locking 
have been observed broadly in a variety of systems
either in 
velocity-force curves and voltage-current curves \cite{Golubov2004}.
Shapiro steps
are hallmarks of
non-Ohmic behavior in voltage-current curves,
where these steps are generated
by the tendency of the response of a system
to be flat across a range of applied driving forces or voltages,
which can be probed with applied AC forces.

%  atoms on surfaces.
%  voltage driven superconducting vortices
%  confined in a Josephson junction,
%  and ac and dc driven
%  charge and spin density waves.
%
%
%mini review of synchronization in electronics
%careful - the following is all simulations
%Recent studies demonstrate
%that coupled electronic oscillators can be
%programmed for use as microcontrollers for robotic locomotion \cite{Dutta2019}.

Recent experiments of particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking in colloids interacting via magnetic dipole
or electrostatic interactions
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a particularly useful medium for studying complex dynamical behaviors.  %cite
Colloid particles are micron-sized,
a relatively large size to control and image.
The interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles \cite{Grier2003}.
This ease in control of colloids has lead to a rich array of 
experimental results  
considered models
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}

%-----------------------------------------------------------
%particle interaction types
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

@@

%describe types of physical systems - particle types
In this work,
we perform 
numerical simulations of confined, driven particles
to model a variety of physical phenomena.
%----------------------------------------------------------------------------
%paper outline
In the following paper we describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

@@

%  Coupling between two singing wineglasses
%Tal Arane, Ana K. R. Musalem and Moti Fridman
%American Journal of Physics 77, 1066 (2009); https://doi.org/10.1119/1.3119175

@@

    \bibitem{Tong1997} E. Y. C. Tong, Lissajous figures The Physics Teacher {\bf 35}, 491 (1997).

@@

@@

    %  \bibitem{} Reichhardt, C., & Reichhardt, C. J. O. (2015). Shapiro steps for skyrmion motion on a washboard potential with longitudinal and transverse ac drives. Physical Review B - Condensed Matter and Materials Physics, 92(22). https://doi.org/10.1103/PhysRevB.92.224432

@@

>>>>

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to interaction or an external stimulus \cite{Pikovsky2003}.
Synchronous behaviors where oscillators move in-phase (i.e. in time)
are observed
in many everyday systems
such as the
flickering candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronzation such as 
birds coordinating wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternating croaking patterns \cite{Aihara2014},
and humans synchronizing clapping in time with music \cite{Tranchant2016}.
At a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two pendula on wall-mounted clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms couple
so the frequency ratio or mode 
is an integer number \cite{Bak1986}.
%Dynamical mode locking
The phenomena 
is often studied with simplified computational models
and can be visualized with %demonstrated
with phase plots or Lissajous figures.
A mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
({\it Could these for Fig. 1})

External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance 
%pushing a child on a swing at the high point on her path, or
using pacemaker (an electrical generator)
pulses to regulate a heart beat
or using a pulsed light to control the pattern of flashing
fireflies \cite{Agrawal2013}.
These dynamical drive studies can be performed
on coupled oscillators %\cite{Kuramoto193?}
or a single particle driven across a periodic potential landscape.
A single particle particle oscillator
within a potential well is 
much like a skateboarder in a half-pipe.  % \cite{}.
%Single particle models avoid many-body effects
%while offering insight.
The applied driving force can be constant (dc) or sinusoidal (ac).
With a pure 
dc drive the landscape modulates 
the average particle velocity.
An ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

%Oscillators ordered into chains or lattices.
%has many important
%technological applications
Dynamical mode-locking
is %often
observed
in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}.
A single junction contains 
two superconductoring layers which sandwich an insulating layer.
When subject to an external voltage,
Cooper pairs in the superconducting materials
tunnel through the insulating layer.
Phase-locking is observed as 
%phase-locked current loops \cite{} 
stepped regions in current-voltage (I-V) relationship in these devices,
where voltage is the analog of external driving force
and current is that of particle velocity.
%where a constant current results from a range of voltages
%due to the quantized nature of the tunneling probability [NOT QUITE RIGHT...].
%instead of ohmic relationships.
Known as Shapiro steps, % \cite{},
these phase locked currents  
have been observed due to applied ac voltages in 
single Josephson junctions \cite{Shapiro1963, Golubov2004} and
coupled arrays of junctions \cite{Benz1990}.
%either in 
%velocity-force curves and voltage-current curves \cite{Golubov2004}.
Shapiro steps vary in width depending on the strength of the
applied AC forces,
and are observed in a variety of systems
displaying
%are hallmarks of
non-Ohmic behavior in voltage-current curves,
including
ac and dc driven
charge and spin density waves.

%where these steps are generated
%by the tendency of the response of a system
%to be flat across a range of applied driving forces or voltages,
%which can be probed with
%  atoms on surfaces.
%  voltage driven superconducting vortices
%  confined in a Josephson junction,

%
%
%mini review of synchronization in electronics
%careful - the following is all simulations
%Recent studies demonstrate
%that coupled electronic oscillators can be
%programmed for use as microcontrollers for robotic locomotion \cite{Dutta2019}.

Experiments of colloid particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of mode-locking,
which cannot be easily imaged in electronics.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a useful medium for studying complex dynamical behaviors.  %cite
Colloid particles are micron-sized,
a relatively large size to control and image.
The interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the colloidal particles.
Because colloids are easy to control and study, 
they are used as experimental models
for systems relatively hard to access and visualize,
such as cold atoms or electron gases \cite{Grier2003}. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Mode-locking 
in colloids has been acheived in systems interacting via magnetic dipoles 
or electric forces 
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.
%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [mor%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].

@@

%describe types of physical systems - particle types
In this work,
we perform 
numerical simulations of confined, driven particles
to model a variety of physical phenomena.
%----------------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
In Section~\ref{sec:results} we summarize
our results,
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

\bibitem{Arane2009} T. Arane, A. K. R. Musalem and M. Fridman, Coupling between two singing wineglasses, Am. J. Phys. {\bf 77}, 1066 (2009). %; https://doi.org/10.1119/1.3119175

@@

@@

    \bibitem{Tong1997} E. Y. C. Tong, Lissajous figures, The Physics Teacher {\bf 35}, 491 (1997).

    \bibitem{Agrawal2013} D. K. Agrawal, J. Woodhouse, and A. A. Seshia, Physical Review Letters {\bf 111}, 084101 (2013).

@@

    \bibitem{Benz1990}  S. P. Benz, M. S. Rzchowski, M. Tinkham, and C. J. Lobb, Fractional giant Shapiro steps and spatially correlated phase motion in 2D Josephson arrays, Phys. Rev. Lett. 64, 693 (1990); D. Dom´ınguez and J. V. Jos´e, Giant Shapiro steps with screening currents, Phys. Rev. Lett. 69,
514 (1992).

@@

    \bibitem{Reichhardt2015} C. Reichhardt, and C. J. O. Reichhardt,  Shapiro steps for skyrmion motion on a washboard potential with longitudinal and transverse ac drives. Phys. Rev. B {\bf 92}, (22). (2015).

@@

      \bibitem{Grier2003} D. G. Grier, A revolution in optical manipulation. Nature {\bf 424}, 810–816 (2003).
